Fast, Scalable Phrase-Based SMT Decoding

### 2016-02-03 acl.2016/acl2016.tex
acl paper

We trained a phrase-based system using the Moses toolkit with standard settings. The training data consisted of most of the publicly available Arabic-English data from Opus (Jörg Tiedemann, 2012,) containing over 69 million parallel sentences, and tuned on a held out set. The phrase-table was then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$. All models files were then binarized; the language models were binaized using KenLM (???), the phrase table using Probing PT (???), lexicalized reordering model using the compact datastructure described in ???. These binary formats were choosen for their best-of-class multithreaded performance. Table~\ref{tab:model-files} gives details of the resultant sizes of the model files. For verification with a different dataset, we also used a second system trained on the French-English Europarl corpus. 
\begin{table}[h]
\begin{center}
\begin{tabular}{|l|c|c|} \hline
		& ar-en	& fr-en \\ \hline
Phrase table  	& 17 	& 5.8 \\
Language model 	& 3.1  	& 1.8 \\ 
Lex-re model	& 2.3	& 637MB \\ \hline
\end{tabular}
\end{center}
\caption{Model sizes in GB}
\label{tab:model-files}
\end{table}
For testing decoding speed, we used a subset of the training data, Table~\ref{tab:test-sets}. The two test set have differing characterics that we are interested in analyzing, ar-en have short sentences while fr-en have overly long sentences.

@@

Where we need to compare the model score of the algorithms, we used a held out set; ??? for ar-en and ??? for fr-en.

@@

Over 24\% of the Moses decoder running time is spent on memory management and this increases when more threads are used, Table~\ref{tab:optimizing-memory}, braking the scaling of the decoder to many cores. By contrast, our decoder spends 11\% on memory management and does not significantly increase with more cores.
\begin{table}[h]
\begin{center}
\begin{tabular}{|l|c|c|c|c|} \hline
		& \multicolumn{2}{c|} {Moses}	& \multicolumn{2}{c|} {Our Work} \\ \hline
\# threads	& 1 		& 32	& 1 		& 32  \\ \hline
Memory  	& 24		& 39 	& 11		& 13 \\
LM 		& 12	 	& 2 	& 47		& 38 \\ 
Phrase-table	& 9	 	& 5 	& 2		& 4 \\ 
Lex RO 		& 8	 	& 2 	& 2		& 2 \\ 
Search 		& 2	 	& 0 	& 14		& 19 \\ 
Misc/Unknown	& 45	 	& 39 	& 24		& 29 \\ \hline
\end{tabular}
\end{center}
\caption{Profile of \%age decoding time}
\label{tab:optimizing-memory}
\end{table}

@@

We also note that distortion limit which constrains hypothesis extension is dependent on the hypothesis' coverage vector, $C$ \emph{and} the end position of most recent source word that has been translated, $e$. The distortion limit must be checked for every instance of a hypothesis and translation rule, Figure~\ref{algo:Hypothesis Expansion with Unsorted Stack}.
\begin{figure} [h]
\begin{algorithmic}
%\REQUIRE source sentence $s$, translation options
\FORALL{ $hypo$ in $stack_{|C|}$} 
  \FORALL{ translation rules}
      \IF {can-expand($C(hypo)$, $e(hypo)$, translation rule range)}
	\STATE expand hypo with translation rule $\rightarrow$ new hypo
	\STATE add new hypo to next stack
      \ENDIF
  \ENDFOR % { translation options}
\ENDFOR % { hypothesis in $stack_i$} 
\end{algorithmic}
\caption{Hypothesis Expansion with Unsorted Stack}
\label{algo:Hypothesis Expansion with Unsorted Stack}
\end{figure}
However, by separating hypotheses into \emph{'ministacks'} according to coverage and end position, the distortion limit only needs to be checked for each ministack, Figure~\ref{algo:Hypothesis Expansion with Sorted Stack}.
\begin{figure} [h]
\begin{algorithmic}
%\REQUIRE source sentence $s$, translation options
\FORALL{ $ministack_{C, e}$ in $stack_{|C|}$} 
  \FORALL{ translation rules}
    \IF {can-expand($C$, $e$, translation rule range)}
      \FORALL{ $hypo$ in $ministack_{C, e}$} 
	\STATE expand hypo with translation rule $\rightarrow$ new hypo
	\STATE add new hypo to next ministack
      \ENDFOR
    \ENDIF
  \ENDFOR % { translation options}
\ENDFOR % { hypothesis in $stack_i$} 
\end{algorithmic}
\caption{Hypothesis Expansion with Sorted Stack}
\label{algo:Hypothesis Expansion with Sorted Stack}
\end{figure}
We therefore looked at the effects of three stack configurations.
\begin{enumerate}
  \item coverage cardinality,
  \item coverage,
  \item coverage and end position of most recent translated source word.
\end{enumerate}
We therefore looked at the effects of three stack configurations.
\begin{enumerate}
  \item coverage cardinality,
  \item coverage,
  \item coverage and end position of most recent translated source word.
\end{enumerate}
Table~\ref{tab:stack-configuration} present the decoding time (with 32 threads) for each stack configuration, and the average model on the held out test set.
\begin{table}[h]
\begin{center}
\begin{tabular}{|l|p{17mm}|p{17mm}|} \hline
			& Decoding \newline Time(secs)	& Avg model \newline score \\ \hline
Coverage cardinality  	& 186 	& -8.59922 \\
Coverage 		& ??? 	& ??? \\ 
Coverage \& end pos	& 196	& -8.59922 \\ \hline
\end{tabular}
\end{center}
\caption{Decoding time and average model scores for different stack configurations}
\label{tab:stack-configuration}
\end{table}

>>>>

We trained a phrase-based system using the Moses toolkit with standard settings. The training data consisted of most of the publicly available Arabic-English data from Opus (Jörg Tiedemann, 2012,) containing over 69 million parallel sentences, and tuned on a held out set. The phrase-table was then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$. All models files were then binarized; the language models were binaized using KenLM (???), the phrase table using Probing PT (???), lexicalized reordering model using the compact datastructure described in ???. These binary formats were choosen for their best-in-class multithreaded performance. Table~\ref{tab:model-files} gives details of the resultant sizes of the model files. For verification with a different dataset, we also occasionally used a second system trained on the French-English Europarl corpus (2m parallel sentences). 
\begin{table}[h]
\begin{center}
\begin{tabular}{|l|c|c|} \hline
		& ar-en	& fr-en \\ \hline
Phrase table  	& 17 	& 5.8 \\
Language model 	& 3.1  	& 1.8 \\ 
Lex-re model	& 2.3	& 637MB \\ \hline
\end{tabular}
\end{center}
\caption{Model sizes in GB}
\label{tab:model-files}
\end{table}
For testing decoding speed, we used a subset of the training data, Table~\ref{tab:test-sets}. The two test scenarios have differing characterics that we are interested in analyzing, ar-en have short sentences with large models while fr-en have overly long sentences with smaller models.

@@

Where we need to compare model scores, we used held out test sets; a 2,000 sentence subset of OpenSubtitles for ar-en and the 3,003 sentence newstest2011 test set for fr-en.

@@

Over 24\% of the Moses decoder running time is spent on memory management and this increases when more threads are used, Table~\ref{tab:optimizing-memory}, dampening the scalability ofthe decoder. By contrast, our decoder spends 11\% on memory management and does not significantly increase with more cores.
\begin{table}[h]
\begin{center}
\begin{tabular}{|l|c|c|c|c|} \hline
		& \multicolumn{2}{c|} {Moses}	& \multicolumn{2}{c|} {Our Work} \\ \hline
\# threads	& 1 		& 32	& 1 		& 32  \\ \hline
Memory  	& 24		& 39 	& 11		& 13 \\
LM 		& 12	 	& 2 	& 47		& 38 \\ 
Phrase-table	& 9	 	& 5 	& 2		& 4 \\ 
Lex RO 		& 8	 	& 2 	& 2		& 2 \\ 
Search 		& 2	 	& 0 	& 14		& 19 \\ 
Misc/Unknown	& 45	 	& 39 	& 24		& 29 \\ \hline
\end{tabular}
\end{center}
\caption{Profile of \%age decoding time}
\label{tab:optimizing-memory}
\end{table}

@@

We also note that distortion limit which constrains hypothesis extension is dependent on the hypothesis' coverage vector, $C$ \emph{and} the end position of most recent source word that has been translated, $e$. The distortion limit must be checked for every instance of a hypothesis and translation rule, Figure~\ref{algo:Hypothesis Expansion with Unsorted Stack}.
\begin{figure} [h]
\begin{algorithmic}
%\REQUIRE source sentence $s$, translation options
\FORALL{ $hypo$ in $stack_{|C|}$} 
  \FORALL{ translation rules}
      \IF {can-expand($C(hypo)$, $e(hypo)$, translation rule range)}
	\STATE expand hypo with translation rule $\rightarrow$ new hypo
	\STATE add new hypo to next stack
      \ENDIF
  \ENDFOR % { translation options}
\ENDFOR % { hypothesis in $stack_i$} 
\end{algorithmic}
\caption{Hypothesis Expansion with Unsorted Stack}
\label{algo:Hypothesis Expansion with Unsorted Stack}
\end{figure}
However, by separating hypotheses into \emph{'ministacks'} according to coverage and end position, the distortion limit only needs to be checked for each ministack, Figure~\ref{algo:Hypothesis Expansion with Sorted Stack}.
\begin{figure} [h]
\begin{algorithmic}
%\REQUIRE source sentence $s$, translation options
\FORALL{ $ministack_{C, e}$ in $stack_{|C|}$} 
  \FORALL{ translation rules}
    \IF {can-expand($C$, $e$, translation rule range)}
      \FORALL{ $hypo$ in $ministack_{C, e}$} 
	\STATE expand hypo with translation rule $\rightarrow$ new hypo
	\STATE add new hypo to next ministack
      \ENDFOR
    \ENDIF
  \ENDFOR % { translation options}
\ENDFOR % { hypothesis in $stack_i$} 
\end{algorithmic}
\caption{Hypothesis Expansion with Sorted Stack}
\label{algo:Hypothesis Expansion with Sorted Stack}
\end{figure}
We therefore looked at the effects of three stack configurations.
\begin{enumerate}
  \item coverage cardinality,
  \item coverage,
  \item coverage and end position of most recent translated source word.
\end{enumerate}
We therefore looked at the effects of three stack configurations.
\begin{enumerate}
  \item coverage cardinality,
  \item coverage,
  \item coverage and end position of most recent translated source word.
\end{enumerate}
Table~\ref{tab:stack-configuration-model-score} present the average model on the held out test set, while Table~\ref{tab:stack-configuration-time} presents the decoding time.
\begin{table}[h]
\small
\begin{center}
\begin{tabular}{|r|p{12mm}|p{12mm}|p{13mm}|} \hline
Pop-limit	& Coverage \newline cardinality	& Coverage	& Coverage \newline \& end pos \\ \hline
100  		& -8.64513 			& ???	& -8.64513 \\
500		& -8.59563			& ??? 	& -8.59612 \\
1,000	 	& -8.5870			& ???	& -8.5870 \\ 
5,000 		& -8.58165			& ???	& -8.58098 \\ 
10,000 		& -8.58133			& ???	& -8.58015 \\ \hline
\end{tabular}
\end{center}
\caption{Average model scores for different stack configurations}
\label{tab:stack-configuration-model-score}
\end{table}

\begin{table}[h]
\small
\begin{center}
\begin{tabular}{|r|p{12mm}|p{12mm}|p{13mm}|} \hline
Pop-limit	& Coverage \newline cardinality	& Coverage	& Coverage \newline \& end pos \\ \hline
100  		& 73 			& ???	& 72 \\
500		& 237			& ??? 	& 229 \\
1,000	 	& 416			& ???	& 423 \\ 
5,000 		& 1,930			& ???	& 2153 \\ 
10,000 		& 			& ???	& 4576 \\ \hline
\end{tabular}
\end{center}
\caption{Decoding time with 32-threads for different stack configurations (sec)}
\label{tab:stack-configuration-time}
\end{table}
